Multisymplectic observables and higher Courant algebroids

### 2024-02-20 msgauge-Caprarola.tex
Lento incedere - caprarola Talk.

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}[fragile]{A \emph{pentagonal} diagram in symplectic geometry}
	Given a \alert{symplectic mfd.} $(M,\omega)$ ...
	\vfill
	\begin{center}
		\includestandalone[width=.95\textwidth]{Pictures/Frame_Embedding_Diagram_symplectic}
	\end{center}
	\vfill
	\begin{minipage}[t][8.5em][t]{\textwidth}
		\begin{itemize}
			\only<1-3>{
			\item<1-> \alert<1>{... there is a naturally associated Poisson algebra ...}
			\item<2-> \alert<2>{... and also a (standard twisted) Lie Algebroid}.
			\item<3-> A Lie algebroid is a "controlled" $\infty$-dimensional Lie algebra given (in this case) by
			\begin{displaymath}
					\left[\binom{x_1}{f_1},\binom{x_2}{f_2}\right]
					~=~
					\binom{[x_1,x_2]}{x_1(f_2)-x_2(f_1)-\omega(x_1,x_2)}
				\end{displaymath}
			}
			\item[]<4->
				\quad\\
				\begin{thmblock}[There exists an embedding of Lie algebras.]
					\begin{displaymath}
						\begin{tikzcd}
							\Psi~:&[-1em] C^{\infty}(M)_\omega \ar[r,"\Psi"]& \Gamma(TM\oplus \mathbb{R})_\omega
							\\[-2em]
							& f \ar[r,mapsto] & \binom{\mathscr{v}_f}{f}
						\end{tikzcd}		
					\end{displaymath}
				\end{thmblock}
		\end{itemize}
	\end{minipage}
\end{frame}
\note{}
%-------------------------------------------------------------------------------------------------------------------------------------------------

@@

\begin{frame}[t, fragile]{Multisymplectic manifolds} %Fragile -->workaround tikzcd
	\begin{center}
		$-$ \emph{multisymplectic means \textbf{going higher} in the degree of $\omega$} $-$
	\end{center}
	\pause
	\begin{defblock}[$n$-plectic manifold ~\emph{(Cantrijn, Ibort, De Le\'on)} \cite{Cantrun2017}]
		\includestandalone[width=0.95\textwidth]{Pictures/Figure_multisym}	
	\end{defblock}
	%
	\vfill
	%
	%
	\pause
	\begin{block}{Examples:}
		\begin{itemize}
			\item[$\bullet$] 1-plectic $=$ symplectic
			\item[$\bullet$] Any oriented $(n+1)$-dimensional manifold is $n$-plectic w.r.t. the volume form.
			\item[$\bullet$] The multicotangent bundle $\Lambda^n T^\ast Q$ is naturally $n$-plectic.
		\end{itemize}
	\end{block}			 
%
	\pause
	\begin{block}{Historical motivation}
		Mechanics: geometrical foundations of \textit{(first-order)} field theories.
		\begin{itemize}
		 \item[•] Kijowski, W. Tulczyjew \cite{Kijowski1979}; %(1979)
		 \item[•] Cariñena, Crampin, Ibort \cite{Carinena1991b};% (1991)
		 \item[•] Gotay, Isenberg, Marsden, Montgomery \cite{Gimmsy1};%(1998)
		 \\ $\cdots$
		\end{itemize}
	\end{block}
\end{frame}

@@

\begin{frame}{Hamiltonian forms}
	\only<1>{
	\medskip
	\begin{columns}[T]
		\setlength{\belowdisplayskip}{5pt}
		\begin{column}{.50\linewidth}
			\centering \it
			$-$ symplectic case $-$
			\vspace{20em}		
		\end{column}	
		%
		\onslide{\vrule{}}
		%
		\begin{column}{.50\linewidth}
			\centering \it
			$-$ $n$-plectic case $-$
			\vspace{20em}		
		\end{column}	
	\end{columns}
	}
	\pause
	%
	\begin{defblock}[Hamiltonian $(n-1)$-forms]
		\begin{displaymath}
			\Omega^{n-1}_{ham}(M,\omega) 	:=
			\biggr\{ \sigma \in  \Omega^{n-1}(M) \; \biggr\vert \; 
				\exists \vHam_\sigma \in \mathfrak{X}(M) ~:~ 
				\tikz[baseline,remember picture]{\node[rounded corners,
                        fill=orange!5,draw=orange!30,anchor=base]            
            			(target) {$d \sigma = -\iota_{\vHam_\sigma} \omega$ };
            	}				
				~\biggr\} 
			\end{displaymath}
	\end{defblock}
	%
	\onslide<2>{
		\tikz[overlay,remember picture]
		{
			\node[rounded corners,
                 fill=orange!5,draw=orange!30,anchor=base]
            	 (base) at ($(current page.north east)-(2,1)$) [rotate=-0,text width=3.5cm,align=center] {\footnotesize{\textcolor{red}{Hamilton-DeDonder-Weyl \\equation}}};
		}	
		\begin{tikzpicture}[overlay,remember picture]
		    	\path[->] (base.south east) edge[bend left,red](target.east);
	    \end{tikzpicture}
	}
	%
	\vspace{-1em}
	\begin{columns}[T]
		\setlength{\belowdisplayskip}{5pt}
		\begin{column}{.50\linewidth}
			%
			\centering \it
			$-$ symplectic case $-$
			\onslide<3->{
			\begin{thmblock}[Observables Poisson algebra]
				$C^\infty(M,\omega)$ endowed with
				\vspace{-.5em}
				\begin{displaymath}
					\lbrace \sigma_1, \sigma_2 \rbrace =			
					~ - \iota_{\vHam_1}\iota_{\vHam_2} \omega 
					~= \mathcal{L}_{\vHam_1} \sigma_2
				\end{displaymath}			
				forms a Poisson algebra.
			\end{thmblock}
			}
			%
			\onslide<4->{
			\vspace{1em}
			\begin{itemize}
				\item[\cmark] Skew-symmetric;
				\item[\cmark] multiplication of observables;
				\item[\cmark] Leibniz Rule;
				\item[\cmark] Jacobi equation;
			\end{itemize}		
			}		
		\end{column}	
		%
		\onslide<1->{\vrule{}}
		%
		\begin{column}{.50\linewidth}
			\centering \it
			$-$ $n$-plectic case $-$
			\onslide<5->{			
			\begin{thmblock}[Observables $L_\infty$-algebra]
				$\Omega^{n-1}_{ham}(M,\omega)$ endowed with
				\vspace{-.5em}
				\begin{displaymath}
					\lbrace \sigma_1, \sigma_2 \rbrace =			
					~ - \iota_{\vHam_1}\iota_{\vHam_2} \omega 
				\end{displaymath}			
				can be extended to a \\ $L_\infty-algebra$.
			\end{thmblock}
			}
			%
			\onslide<6->{
			\begin{itemize}
				\item[\cmark] Skew-symmetric;
				\item[\xmark] multiplication of observables;
				\item[\xmark] Jacobi equation;
				%\\ \hspace*{4.25em} full-fledged Jacobi equation;
				\item[\smark] Jacobi equation \emph{up to homotopies}.
			\end{itemize}			
			}
		\end{column}	
	\end{columns}
\end{frame}
%-------------------------------------------------------------------------------------------------------------------------------------------------

@@

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}[fragile]{Lie $\infty$-algebra of Observables (higher observables) }
	Let be $(M,\omega)$ a $n$-plectic manifold.
	  	\vfill
	\begin{defblock}[$L_\infty$-algebra of observables ~\emph{(Rogers)} \cite{Rogers2010}]
		\medskip
		\hspace{.25em} Is a cochain-complex $(L,\{\cdot\}_1)$ \\
		\vspace{-1em}
		\begin{center}
			\includestandalone{Pictures/hyb-Frame_Observables}
		\end{center}
		\onslide<2->{
			\bigskip
			\hspace{.25em} with $n$ (skew-symmetric) multibrackets $(2 \leq k \leq n+1)$\\
			\vspace{-1em}
			\onslide<3->{
				\begin{center}
					\includestandalone{Pictures/Equation_Multibracket}	
				\end{center}
			}
			\medskip
		}
		%
	\end{defblock}
%	\onslide<3->{
%		\emph{Higher analogue} of the \emph{Poisson algebra structure} associated to a symplectic mfd.
%	\vfill
%	\begin{columns}
%		\hfill
%		\begin{column}{.11\linewidth}	
%			If $n>1$:
%			
%		\end{column}	
%		\begin{column}{.8\linewidth}
%		\begin{itemize}
%			\item[\xmark] \textcolor{red}{we lose} :\quad multiplication of observables, Jacobi equation;
%			%\\ \hspace*{4.25em} full-fledged Jacobi equation;
%			\item[\cmark] \textcolor{green}{we gain} :\quad brackets with arities different than two,\\
%			\hspace*{4.25em}
%			 Jacobi equation \emph{up to homotopies}.
%		\end{itemize}		
%		\end{column}		
%	\end{columns}
%	}
  \end{frame}
%-------------------------------------------------------------------------------------------------------------------------------------------------

@@

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}[t]{Vinogradov Algebroids}
	\begin{defblock}[Vinogradov algebroid (higher Courant)]
		\includestandalone[width=0.95\textwidth]{Pictures/hyb-Frame_vinogradov}		
	\end{defblock}
	\vfill
	\begin{itemize}
		\item<7-> $(n=1) ~ \Rightarrow$ standard twisted \emph{Lie algebroid};
		\item<8-> $(n=2) ~ \Rightarrow$ standard twisted \emph{Courant algebroid};
	\end{itemize}

@@

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}[fragile]{Vinogradov $L_\infty$-algebra}
	Vin. alg.oids 
	$\quad\simeq\quad$ 	
	$NQ$-manifolds {\small($L_\infty$-algebroids)}
	$\quad\Rightarrow\quad$ 
	{\small Associated} $L_\infty$-algebra
	%
	\pause
	\begin{defblock}[Vinogradov $L_\infty$-algebra (Zambon) \cite{Zambon2012}]
		\includestandalone[width=0.95\textwidth]{Pictures/hyb-Frame_vinogradov-Linfty}	
	\end{defblock}
	\vfill
	\only<5->{
		\tikz[overlay,remember picture]
		{
			\node[rounded corners,
                 fill=gray!1,draw=gray!30,anchor=base]            
            	 (base) at ($(current page.south)+(0,.5)$) [rotate=-0,text width=10cm,align=center] { \footnotesize{\color{gray}{
            	 $e_i = \binom{X_i}{\alpha_i} \in \mathfrak{X}(M)\oplus \Omega^{n-1}(M)$ 
            	 \quad~,\qquad
            	 $f_i \in \bigoplus_{k=0}^{n-2}\Omega^k(M)$.
            	 }}};
		}			
	}

@@

%-------------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Does Rogers $L_\infty$ embeds into Vinogradov?}
\subcheckpoint	
%-------------------------------------------------------------------------------------------------------------------------------------------------
%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}[fragile,t]{Embedding observables $L_\infty$-algebra into Vinogradov $L_\infty$-algebra}
	Consider now $\omega$ \alert{$n$-plectic}~.
	\vfill
	\only<4->{\vspace{-5em}}
	\begin{center}
		\includestandalone[width=.8\textwidth]{Pictures/hyb-Frame_Embedding_Diagram_k-plectic}
	\end{center}
	\vfill
	\begin{minipage}[t][8.5em][t]{\textwidth}
	\only<4->{
	\begin{thmblock}[Embedding of $L_\infty$-algebras  $\Psi:L_\infty(M,\omega)\hookrightarrow L_{\infty}(E^n,\omega)$\quad \cite{Miti2022}.]
	\begin{itemize}[leftmargin=0pt]
		\item[$\cdot$]<5-> 
			consider the graded vector subspace $\mathcal{A}$
			\onslide<6->{
				\begin{displaymath}
				\mathclap{
				{\mathcal{A}^k} =
				\begin{cases}
			\left.\left\lbrace
			\binom{X}{\alpha}\in \mathfrak{X}(M)\oplus \Omega^{n-1}(M)
			~ \right\vert ~
			\iota_X \omega = -d \alpha\right\rbrace
	&\quad k=0,\\
					\Omega^{n-1+k}(M) &\quad -n+1 \leq k < 0.
				\end{cases}			
				}
				\end{displaymath}	
			}					
		\item[$\cdot$]<6-> 
			restrict the two $L_\infty$-structures to $\pi$ and $\mu$ on $\mathcal{A}$
		\item[$\cdot$]<7->

@@

\begin{frame}[fragile]{Multisymplectic manifolds} %Fragile -->workaround tikzcd
	\begin{defblock}[$n$-plectic manifold ~\emph{(Cantrijn, Ibort, De Le\'on)} \cite{Cantrun2017}]
		\includestandalone[width=0.95\textwidth]{Pictures/Figure_multisym}	
	\end{defblock}
	%
	\begin{defblock}[Non-degenerate $(n+1)$-form]
		\begin{columns}
			\begin{column}{.45\linewidth}
				\centering{
				The $\omega^\flat$ (flat) bundle map is injective.
				}
			\end{column}
			\begin{column}{.5\linewidth}
						\vspace{-.5em}
				\[
				\begin{tikzcd}[column sep= small,row sep=0ex,
				/tikz/column 1/.append style={anchor=base east}]
				    \omega^\flat \colon T M \ar[r]& \Lambda^n T^\ast M \\
  						 (x,u) \ar[r, mapsto]& (x,\iota_{u} \omega_x)						
				\end{tikzcd}	
				\]
			\end{column}
		\end{columns}
	\end{defblock}
	%
	\vfill
	%
	\begin{block}{Examples:}
		\vspace{-.5em}
		\setbeamercovered{transparent}
		\begin{itemize}[<+->]
			\item[$\bullet$] $n=1$ \qquad\qquad\qquad $\Rightarrow$\quad $\omega$ is a symplectic form
			\item[$\bullet$]  $n=(dim(M)-1)$ \quad$\Rightarrow$\quad $\omega$ is a volume form
			%Any oriented $(n+1)$-dimensional manifold is $n$-plectic w.r.t. the volume form.
			%\item[$\bullet$] Let $G$ a semisimple Lie group and $\langle\cdot,\cdot \rangle$  its killing form. Then $\langle [\cdot,\cdot],\cdot \rangle$ extends to a biinvariant multisymplectic form $\omega$.
			\item[$\bullet$] Let $Q$ a smooth manifold, the multicotangent bundle $\Lambda^n T^\ast Q$ is naturally $n$-plectic.%
			\quad
			\textit{(cfr, \href{https://arxiv.org/abs/physics/9801019}{GIMMSY} construction for classical field theories)}
		\end{itemize}
	\end{block}

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}{From symplectic to {multi}symplectic} 
	%
	\begin{center}
		$-$ \emph{multisymplectic means \textbf{going higher} in the degree of $\omega$} $-$
	\end{center}
	\pause
	\begin{defblock}[$n$-plectic manifold ~\emph{(Cantrijn, Ibort, De Le\'on)} \cite{Cantrun2017}]
		\includestandalone[width=0.95\textwidth]{Pictures/Figure_multisym}	
	\end{defblock}
	%
	\pause
		\begin{table}
			\begin{tabular}{c c c}
				symplectic forms \small($n=1$) & $\leftrightsquigarrow$ & volume forms \small($n= \text{dim}(M)-1$)
			\end{tabular}
		\end{table}

	\vfill
	\pause
	\begin{block}{Historical motivation}
		Mechanics: geometrical foundations of \textit{(first-order)} field theories.
	\end{block}
	%
	\begin{table}
		\ifHandout
		%
		\else
		\only<4>{
		\begin{tabular}{|p{0.2\textwidth}|p{0.3\textwidth}|p{0.35\textwidth}|} 
            \hline
            \parbox[][20pt][c]{0.2\textwidth}{mechanics} & \multicolumn{2}{c|}{geometry} \\
            \hline
            \parbox[][20pt][c]{0.2\textwidth}{phase space} & symplectic manifold &  \\[.25em]
            \parbox[][20pt][c]{0.2\textwidth}{classical \\ observables} & Poisson algebra &  \\[.25em]
            \parbox[][20pt][c]{0.2\textwidth}{symmetries} &  group actions admitting comoment map &  
            \\
            \hline
  \multicolumn{1}{c}{}
            &  \multicolumn{1}{@{}c@{}}{$\underbrace{\hspace*{.3\textwidth}}_{\text{point-like particles systems}}$} 
            &  \multicolumn{1}{@{}c@{}}{}              \\
		\end{tabular}
		}
		\fi
		\onslide<5->{
		\begin{tabular}{|p{0.2\textwidth}|p{0.3\textwidth}|p{0.35\textwidth}|} 
            \hline
            \parbox[][20pt][c]{0.2\textwidth}{mechanics} & \multicolumn{2}{c|}{geometry} \\
            \hline
            \parbox[][20pt][c]{0.2\textwidth}{phase space} & symplectic manifold & multisymplectic manifold \\[.25em]
            \parbox[][20pt][c]{0.2\textwidth}{classical \\ observables} & Poisson algebra & $L_\infty$-algebra \\[.25em]
            \parbox[][20pt][c]{0.2\textwidth}{symmetries} &  group actions admitting comoment map & group actions admitting (homotopy) comomentum map
            \\
            \hline
  			\multicolumn{1}{c}{}
            &  \multicolumn{1}{@{}c@{}}{$\underbrace{\hspace*{.3\textwidth}}_{\text{point-like particles systems}}$} 
            &
            \multicolumn{1}{@{}c@{}}{$\underbrace{\hspace*{.3\textwidth}}_{\text{field-theoretic systems}}$} 
               \\
		\end{tabular}
		}
	\end{table}

\end{frame}
\note[itemize]{
	\item Historically, the interest in multisymplectic manifolds, has been motivated by the need for understanding the geometrical foundations of first-order classical field theories.
	The key point is that, just as one can associate a symplectic manifold to an ordinary classical mechanical system (e.g. a single
point-like particle constrained to some manifold), it is possible to associate a multisymplectic
manifold to any classical field system (e.g. a continuous medium like a filament or a fluid). See frame Extra-\ref{Frame:Ms-Field-Mechanics}

	\item General ideas basic parallelisms with caveats
	\item caveat: points in multiphase spaces are not states
	\item the table hides the duality between geometric and algebraic approaches to the problem.
	\item	Mechanics: geometrical foundations of \textit{(first-order)} field theories.
		\begin{itemize}
		 \item[-] Kijowski, W. Tulczyjew \cite{Kijowski1979}; %(1979)
		 \item[-] Cariñena, Crampin, Ibort \cite{Carinena1991b};% (1991)
		 \item[-] Gotay, Isenberg, Marsden, Montgomery \cite{Gimmsy1};%(1998)
		 \\ $\cdots$
		\end{itemize}
}
%-------------------------------------------------------------------------------------------------------------------------------------------------

>>>>

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}[fragile]{A \emph{pentagonal} diagram in symplectic geometry}
	Given a \alert{symplectic mfd.} $(M,\omega)$ ...
	\vfill
	\begin{center}
		\includestandalone[width=.95\textwidth]{Pictures/Frame_Embedding_Diagram_symplectic}
	\end{center}
	\vfill
	\begin{minipage}[t][8.5em][t]{\textwidth}
		\begin{itemize}
			\only<1-3>{
			\item<1-> \alert<1>{... there is a naturally associated Poisson algebra ...}
			\item<2-> \alert<2>{... and also a (standard twisted) Lie Algebroid}.
			\item<3-> A Lie algebroid is a "controlled" $\infty$-dimensional Lie algebra given (in this case) by
			\begin{displaymath}
					\left[\binom{x_1}{f_1},\binom{x_2}{f_2}\right]_\omega
					~=~
					\binom{[x_1,x_2]}{x_1(f_2)-x_2(f_1)-\omega(x_1,x_2)}
				\end{displaymath}
			}
			\item[]<4->
				\quad\\
				\begin{thmblock}[There exists an embedding of Lie algebras.]
					\begin{displaymath}
						\begin{tikzcd}
							\Psi~:&[-1em] C^{\infty}(M)_\omega \ar[r,"\Psi"]& \Gamma(TM\oplus \mathbb{R})_\omega
							\\[-2em]
							& f \ar[r,mapsto] & \binom{\mathscr{v}_f}{f}
						\end{tikzcd}		
					\end{displaymath}
				\end{thmblock}
		\end{itemize}
	\end{minipage}
\end{frame}
\note{}
%-------------------------------------------------------------------------------------------------------------------------------------------------

@@

\begin{frame}[t, fragile]{Multisymplectic manifolds} %Fragile -->workaround tikzcd
	\begin{center}
		$-$ \emph{multisymplectic means \textbf{going higher} in the degree of $\omega$} $-$
	\end{center}
	\pause
	\begin{defblock}[$n$-plectic manifold ~\emph{(Cantrijn, Ibort, De Le\'on)} \cite{Cantrun2017}]
		\includestandalone[width=0.95\textwidth]{Pictures/Figure_multisym}	
	\end{defblock}
	%
	\pause
	%
	\begin{defblock}[Non-degenerate $(n+1)$-form]
		\begin{columns}
			\begin{column}{.45\linewidth}
				\centering{
				The $\omega^\flat$ (flat) bundle map is injective.
				}
			\end{column}
			\begin{column}{.5\linewidth}
						\vspace{-.5em}
				\[
				\begin{tikzcd}[column sep= small,row sep=0ex,
				/tikz/column 1/.append style={anchor=base east}]
				    \omega^\flat \colon T M \ar[r]& \Lambda^n T^\ast M \\
  						 (x,u) \ar[r, mapsto]& (x,\iota_{u} \omega_x)						
				\end{tikzcd}	
				\]
			\end{column}
		\end{columns}
	\end{defblock}		
	\vfill
	%
	%
	\pause
	\begin{block}{Examples:}
		\begin{itemize}
			\item[$\bullet$] 1-plectic $=$ symplectic
			\item[$\bullet$] Any oriented $(n+1)$-dimensional manifold is $n$-plectic w.r.t. the volume form.
			\item[$\bullet$] The multicotangent bundle $\Lambda^n T^\ast Q$ is naturally $n$-plectic.
		\end{itemize}
	\end{block}			 
%

\end{frame}

\begin{frame}[fragile]{Multisymplectic manifolds - motivations} %Fragile -->workaround tikzcd
	\begin{block}{Historical motivation}
		Mechanics: geometrical foundations of \textit{(first-order)} field theories.
		\begin{itemize}
		 \item[•] Kijowski, W. Tulczyjew \cite{Kijowski1979}; %(1979)
		 \item[•] Cariñena, Crampin, Ibort \cite{Carinena1991b};% (1991)
		 \item[•] Gotay, Isenberg, Marsden, Montgomery \cite{Gimmsy1};%(1998)
		 \\ $\cdots$
		\end{itemize}
	\end{block}
	%
	\pause
	\begin{table}
		\ifHandout
		%
		\else
		\only<2>{
		\begin{tabular}{|p{0.2\textwidth}|p{0.3\textwidth}|p{0.35\textwidth}|} 
            \hline
            \parbox[][20pt][c]{0.2\textwidth}{mechanics} & \multicolumn{2}{c|}{geometry} \\
            \hline
            \parbox[][20pt][c]{0.2\textwidth}{phase space} & symplectic manifold &  \\[.25em]
            \parbox[][20pt][c]{0.2\textwidth}{classical \\ observables} & Poisson algebra &  \\[.25em]
            \parbox[][20pt][c]{0.2\textwidth}{symmetries} &  group actions admitting comoment map &  
            \\
            \hline
  \multicolumn{1}{c}{}
            &  \multicolumn{1}{@{}c@{}}{$\underbrace{\hspace*{.3\textwidth}}_{\text{point-like particles systems}}$} 
            &  \multicolumn{1}{@{}c@{}}{}              \\
		\end{tabular}
		}
		\fi
		\onslide<3->{
		\begin{tabular}{|p{0.2\textwidth}|p{0.3\textwidth}|p{0.35\textwidth}|} 
            \hline
            \parbox[][20pt][c]{0.2\textwidth}{mechanics} & \multicolumn{2}{c|}{geometry} \\
            \hline
            \parbox[][20pt][c]{0.2\textwidth}{phase space} & symplectic manifold & multisymplectic manifold \\[.25em]
            \parbox[][20pt][c]{0.2\textwidth}{classical \\ observables} & Poisson algebra & $L_\infty$-algebra \\[.25em]
            \parbox[][20pt][c]{0.2\textwidth}{symmetries} &  group actions admitting comoment map & group actions admitting (homotopy) comomentum map
            \\
            \hline
  			\multicolumn{1}{c}{}
            &  \multicolumn{1}{@{}c@{}}{$\underbrace{\hspace*{.3\textwidth}}_{\text{point-like particles systems}}$} 
            &
            \multicolumn{1}{@{}c@{}}{$\underbrace{\hspace*{.3\textwidth}}_{\text{field-theoretic systems}}$} 
               \\
		\end{tabular}
		}
	\end{table}

@@

\begin{frame}{Hamiltonian forms}
	\begin{defblock}[Hamiltonian $(n-1)$-forms]
		\begin{displaymath}
			\Omega^{n-1}_{ham}(M,\omega) 	:=
			\biggr\{ \sigma \in  \Omega^{n-1}(M) \; \biggr\vert \; 
				\exists \vHam_\sigma \in \mathfrak{X}(M) ~:~ 
				\tikz[baseline,remember picture]{\node[rounded corners,
                        fill=orange!5,draw=orange!30,anchor=base]            
            			(target) {$d \sigma = -\iota_{\vHam_\sigma} \omega$ };
            	}				
				~\biggr\} 
			\end{displaymath}
	\end{defblock}
	\pause
	%

	%
	\onslide<2>{
		\tikz[overlay,remember picture]
		{
			\node[rounded corners,
                 fill=orange!5,draw=orange!30,anchor=base]
            	 (base) at ($(current page.north east)-(2,1)$) [rotate=-0,text width=3.5cm,align=center] {\footnotesize{\textcolor{red}{Hamilton-DeDonder-Weyl \\equation}}};
		}	
		\begin{tikzpicture}[overlay,remember picture]
		    	\path[->] (base.south east) edge[bend left,red](target.east);
	    \end{tikzpicture}
	}
	%
	\vspace{-1em}
	\begin{columns}[T]
		\setlength{\belowdisplayskip}{5pt}
		\begin{column}{.50\linewidth}
			%
			\centering \it
			$-$ symplectic case $-$
			\onslide<3->{
			\begin{thmblock}[Observables Poisson algebra]
				$C^\infty(M,\omega)$ endowed with
				\vspace{-.5em}
				\begin{displaymath}
					\lbrace \sigma_1, \sigma_2 \rbrace =			
					~ - \iota_{\vHam_1}\iota_{\vHam_2} \omega 
					~= \mathcal{L}_{\vHam_1} \sigma_2
				\end{displaymath}			
				forms a Poisson algebra.
			\end{thmblock}
			}
			%
			\onslide<4->{
			\vspace{1em}
			\begin{itemize}
				\item[\cmark] Skew-symmetric;
				\item[\cmark] multiplication of observables;
				\item[\cmark] Leibniz Rule;
				\item[\cmark] Jacobi equation;
			\end{itemize}		
			}		
		\end{column}	
		%
		\onslide<1->{\vrule{}}
		%
		\begin{column}{.50\linewidth}
			\centering \it
			$-$ $n$-plectic case $-$
			\onslide<5->{			
			\begin{thmblock}[Observables $L_\infty$-algebra]
				$\Omega^{n-1}_{ham}(M,\omega)$ endowed with
				\vspace{-.5em}
				\begin{displaymath}
					\lbrace \sigma_1, \sigma_2 \rbrace =			
					~ - \iota_{\vHam_1}\iota_{\vHam_2} \omega 
				\end{displaymath}			
				can be extended to a \\ $L_\infty-algebra$.
			\end{thmblock}
			}
			%
			\onslide<6->{
			\begin{itemize}
				\item[\cmark] Skew-symmetric;
				\item[\xmark] multiplication of observables;
				\item[\xmark] Jacobi equation;
				%\\ \hspace*{4.25em} full-fledged Jacobi equation;
				\item[\smark] Jacobi equation \emph{up to homotopies}.
			\end{itemize}			
			}
		\end{column}	
	\end{columns}
\end{frame}
%-------------------------------------------------------------------------------------------------------------------------------------------------

@@

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}[fragile]{Lie $\infty$-algebra of Observables (higher observables) }
	Let be $(M,\omega)$ a $n$-plectic manifold.
	  	\vfill
	\begin{defblock}[$L_\infty$-algebra of observables ~\emph{(Rogers)} \cite{Rogers2010}]
		\medskip
		\hspace{.25em} Is a cochain-complex $(L,\{\cdot\}_1)$ \\
		\vspace{-1em}
		\begin{center}
			\includestandalone{Pictures/Frame_Observables}
		\end{center}
		\onslide<2->{
			\bigskip
			\hspace{.25em} with $n$ (skew-symmetric) multibrackets $(2 \leq k \leq n+1)$\\
			\vspace{-1em}
			\onslide<3->{
				\begin{center}
					\includestandalone{Pictures/Equation_Multibracket}	
				\end{center}
			}
			\medskip
		}
		%
	\end{defblock}
%	\onslide<3->{
%		\emph{Higher analogue} of the \emph{Poisson algebra structure} associated to a symplectic mfd.
%	\vfill
%	\begin{columns}
%		\hfill
%		\begin{column}{.11\linewidth}	
%			If $n>1$:
%			
%		\end{column}	
%		\begin{column}{.8\linewidth}
%		\begin{itemize}
%			\item[\xmark] \textcolor{red}{we lose} :\quad multiplication of observables, Jacobi equation;
%			%\\ \hspace*{4.25em} full-fledged Jacobi equation;
%			\item[\cmark] \textcolor{green}{we gain} :\quad brackets with arities different than two,\\
%			\hspace*{4.25em}
%			 Jacobi equation \emph{up to homotopies}.
%		\end{itemize}		
%		\end{column}		
%	\end{columns}
%	}
  \end{frame}
%-------------------------------------------------------------------------------------------------------------------------------------------------

@@

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}[t]{Vinogradov Algebroids}
	\begin{defblock}[Vinogradov algebroid (higher Courant)]
		\includestandalone[width=0.95\textwidth]{Pictures/Frame_vinogradov}		
	\end{defblock}
	\vfill
	\begin{itemize}
		\item<6-> $(n=1) ~ \Rightarrow$ standard twisted \emph{Lie algebroid};
		\item<7-> $(n=2) ~ \Rightarrow$ standard twisted \emph{Courant algebroid};
	\end{itemize}

@@

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}[fragile]{Vinogradov $L_\infty$-algebra}
	Vin. alg.oids 
	$\quad\simeq\quad$ 	
	$NQ$-manifolds {\small($L_\infty$-algebroids)}
	$\quad\Rightarrow\quad$ 
	{\small Associated} $L_\infty$-algebra
	%
	\pause
	\begin{defblock}[Vinogradov $L_\infty$-algebra (Zambon) \cite{Zambon2012}]
		\includestandalone[width=0.95\textwidth]{Pictures/Frame_vinogradov-Linfty}	
	\end{defblock}
	\vfill
	\only<6->{
		\tikz[overlay,remember picture]
		{
			\node[rounded corners,
                 fill=gray!1,draw=gray!30,anchor=base]            
            	 (base) at ($(current page.south)+(0,.5)$) [rotate=-0,text width=10cm,align=center] { \footnotesize{\color{gray}{
            	 $e_i = \binom{X_i}{\alpha_i} \in \mathfrak{X}(M)\oplus \Omega^{n-1}(M)$ 
            	 \quad~,\qquad
            	 $f_i \in \bigoplus_{k=0}^{n-2}\Omega^k(M)$.
            	 }}};
		}			
	}

@@

%-------------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Does Rogers $L_\infty$ embeds into Vinogradov?}
\subcheckpoint	
%-------------------------------------------------------------------------------------------------------------------------------------------------
%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}[fragile,t]{Embedding observables $L_\infty$-algebra into Vinogradov $L_\infty$-algebra}
	Consider now $\omega$ \alert{$n$-plectic}~.
	\vfill
	\only<4->{\vspace{-5em}}
	\begin{center}
		\includestandalone[width=.8\textwidth]{Pictures/hyb-Frame_Embedding_Diagram_k-plectic}
	\end{center}
	\vfill
	\begin{minipage}[t][8.5em][t]{\textwidth}
	\only<4->{
	\begin{thmblock}[Embedding of $L_\infty$-algebras  $\Psi:L_\infty(M,\omega)\hookrightarrow L_{\infty}(E^n,\omega)$\quad \cite{Miti2022}.]
	\begin{itemize}[leftmargin=0pt]
		\item[$\cdot$]<5-> 
			consider the graded vector subspace $\mathcal{A}$
			\onslide<5->{
				\begin{displaymath}
				\mathclap{
				{\mathcal{A}^k} =
				\begin{cases}
			\left.\left\lbrace
			\binom{X}{\alpha}\in \mathfrak{X}(M)\oplus \Omega^{n-1}(M)
			~ \right\vert ~
			\iota_X \omega = -d \alpha\right\rbrace
	&\quad k=0,\\
					\Omega^{n-1+k}(M) &\quad -n+1 \leq k < 0.
				\end{cases}			
				}
				\end{displaymath}	
			}					
		\item[$\cdot$]<6-> 
			restrict the two $L_\infty$-structures to $\pi$ and $\mu$ on $\mathcal{A}$
		\item[$\cdot$]<7->

@@
